---
title: "Pengilley-Alana-ada-homework-2"
author: "Alana Pengilley"
date: "23/03/2021"
output:
  pdf_document: default
  html_document: default
---
## Challenge 1 
---
IMDB Movies 
---
```{r}
#load in packages to use
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(mosaic)
library(purrr)

#load in dataset
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/IMDB-movies.csv"
d <- read_csv(f, col_names = TRUE)
d <- filter(d, startYear %in% 1920:1979, runtimeMinutes >= 60, runtimeMinutes <= 180)
head(d) #first 6 rows to check dataset

d <- d %>% 
  mutate("decade" = floor(startYear/10)*10) #add decade column to dataset
d

#create histograms of run times per decade 
histo_runtimes <- ggplot(data = d, aes(x = runtimeMinutes)) + geom_histogram(binwidth = 5, col = "black", fill = "blue") +
  labs(title = "Run Times per Decade", x = "Run Time (mins)", y = "Count") + facet_wrap(~decade)
histo_runtimes

#population mean and SD for runtimeMins 
(results <- group_by(d, decade) %>%
  summarise(avgTime = mean(runtimeMinutes),sdTime = sd(runtimeMinutes)))


#sample of 100 movies per decade and calculate mean, sd, and se for each group
set.seed(1)
n <- 100
s <- group_by(d, decade) %>% sample_n(size = n, replcae = FALSE) %>%
  summarise(single_sample_size = n(), single_sample_mean = mean(runtimeMinutes),
            single_sample_sd = sd(runtimeMinutes), single_sample_se = sd(runtimeMinutes)/sqrt(n))
s

pop_mean <- results$avgTime
pop_sd <- results$sdTime
pop_se <- pop_sd/sqrt(n)

#compare sample and population stats 
(c <- cbind(pop_mean, pop_sd, pop_se, s))

#generate a sampling distribution of mean of runtimeMins
set.seed(1) 
n <- 100
num_trials <- 1000
samp_dist_stats<- 1:num_trials %>%
  map_dfr(
    ~ group_by(d, decade) %>%
      slice_sample(n = 100)%>%
      summarize(samp_means = mean(runtimeMinutes), samp_sd = sd(runtimeMinutes))
  ) %>% mutate(n = 100)

samp_dist_stats 

#mean and sd for each decade from the sampling distribution 
samp_dist_m <- aggregate(samp_dist_stats$samp_means, list(samp_dist_stats$decade), mean)
samp_dist_sd <- aggregate(samp_dist_stats$samp_sd, list(samp_dist_stats$decade), sd) #this will act as the se for the samp dist

#plot histogram of sample distribution
p <- ggplot(data = samp_dist_stats, aes(x = samp_means)) + geom_histogram(binwidth = 1, col ="black", fill = "red") + facet_wrap(~decade)
p #normal distribution 

#compare standard errors
compare <- data.frame("SE estimate from 1 samp" = s$single_sample_se, "SE calculate form pop SD" = pop_se, "SE estimate from samp dist" = samp_dist_sd[,2]) 
knitr::kable(compare)

```

## Challenge 2 
---
Every morning at the same time, a bee bioloigist goes and sits in a field in the morning and watches for forager bees returning to a hive, counting the number that arrive in a one hour window from 7am to 8am. Based on previous knowledge, she believes that the mean number of foragers that will return in that time is 12, roughly one every 6 minutes.
---

```{r}

#a random variable x has poisson distribution with a mean of 12. 

#P(x<=9)
ppois(q=9, lambda = 12)

#P(x=0)
dpois(x=0, lambda = 12)

#P(x=5)
dpois(x=5, lambda = 12)

#P(x>18)
1 - ppois(q=18, lambda = 12) 
  
#plot relevant Poisson mass function in range 0:24
plotDist("pois", lambda = 12, xlim = c(0,24), kind = "histogram", main = "PMF for Poisson",xlab = "Number of Arriving Bees", ylab = "Frequency")

#simulate 1460 results from this distribution
set.seed(100)
sim <- rpois(n=1460, lambda = 12)

#histogram

p2 <- histogram(sim, xlim = c(0, 24),
  main = "Poisson Distribution of simulated results, lambda = 12",
  type = "density", width = 1, center = 12) 
p2
mean(sim)
var(sim)
#the simulated results are very similar to those displayed on the probability mass function plot above. Both are a mean and variance of approx. 12. 

```

##Challenge 3
---
Zombies data set. This data includes the first and last name and gender of the entire population of 1000 people who have survived the zombie apocalypse and are now ekeing out an existence somewhere on the East Coast, along with several other variables (height, weight, age, number of years of education, number of zombies they have killed, and college major).
---

``` {r}
library(ggplot2)
library(patchwork)
library(gridExtra)
library(radiant)
library(cowplot)
library(purrr)

f <- "https://raw.githubusercontent.com/difiore/ada-2021-datasets/master/zombies.csv"
z <- read_csv(f, col_names = TRUE)

survivors <- select(z, "gender", "height", "weight", "age", "zombies_killed", "years_of_education")

#population mean and SD for each quantitative random variable

pop_mean <- summarise(survivors, mean(height), mean(weight), mean(age), mean(zombies_killed), mean(years_of_education))

pop_sd <- summarise(survivors, sdpop(height), sdpop(weight), sdpop(age), sdpop(zombies_killed), sdpop(years_of_education))

pop_summary<- data.frame(variable=c("Height", "Weight", "Age", "Kills", "Education"), MEAN=t(pop_mean), SD=t(pop_sd))
knitr::kable(pop_summary)

#boxplots of each variable by gender
b1 <- ggplot(survivors, aes(x = gender, y = height)) + geom_boxplot() 
b2 <- ggplot(survivors, aes(x = gender, y = weight)) + geom_boxplot() 
b3 <- ggplot(survivors, aes(x = gender, y = age)) + geom_boxplot() 
b4 <- ggplot(survivors, aes(x = gender, y = zombies_killed)) + geom_boxplot() 
b5 <- ggplot(survivors, aes(x = gender, y = years_of_education)) + geom_boxplot() 
plot_grid(b1, b2, b3, b4, b5, ncol = 3)


#height and weight in relation to age by gender 
p1 <- ggplot(data = survivors, aes(
  x = age,
  y = height,
  color = factor(gender)))
p1 <- p1 + xlab("Height") + ylab("Age")
p1 <- p1 + geom_point(na.rm = TRUE)
p1 <- p1 + theme(legend.position = "bottom", legend.title = element_blank())

p2 <- ggplot(data = survivors, aes(
  x = age,
  y = weight,
  color = factor(gender)))
p2 <- p2 + xlab("Weight") + ylab("Age")
p2 <- p2 + geom_point(na.rm = TRUE)
p2 <- p2 + theme(legend.position = "bottom", legend.title = element_blank())

plot_grid(p1, p2, nrow = 1)
```
---
Do these variables (age, weight and height) seem related? 
Yes, both variables (height and weight) increase with age. 
---

---
Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not?
(normal = height, weight, age), (Possion = kills and education)
---
``` {r}
#Height
par(mfrow= c(1,2))
histogram(survivors$height, main = "Height", xlab = "") 
qqnorm(survivors$height, main = "QQ Plot - Height")
qqline(survivors$height, col = "blue")
#normal 

#Weight
par(mfrow= c(1,2))
histogram(survivors$weight, main = "Weight", xlab = "") 
qqnorm(survivors$weight, main = "QQ Plot - Weight")
qqline(survivors$weight, col = "blue")
#normal

#Age
par(mfrow= c(1,2))
histogram(survivors$age, main = "Age", xlab = "") 
qqnorm(survivors$age, main = "QQ Plot - Age")
qqline(survivors$age, col = "blue") 
#normal

#Number of zombies killed 
par(mfrow= c(1,2))
histogram(survivors$zombies_killed, main = "Kills", xlab = "") 
qqnorm(survivors$zombies_killed, main = "QQ Plot - Kills")
qqline(survivors$zombies_killed, col = "blue") 
#Poisson

#Years of education
par(mfrow= c(1,2))
histogram(survivors$years_of_education, main = "Education", xlab = "") 
qqnorm(survivors$years_of_education, main = "QQ Plot - Education")
qqline(survivors$years_of_education, col = "blue") 
#Poisson
```
---
Sample ONE subset of 50 zombie apocalypse survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable based on this sample and use that to construct a 95% confidence interval for each mean.
---

```{r}
#sample of 50 survivors 
set.seed(1)
n <- 50 
s <- sample_n(survivors, size = n, replace = FALSE)

#means
s_means <- summarise(s, across(c(height, weight, age, zombies_killed, years_of_education), list(mean=mean)))
#sds
s_sd <- summarise(s, height_sd = sd(height),weight_sd = sd(weight),age_sd = sd(age),kills_sd = sd(zombies_killed),edu_sd = sd(years_of_education))
#(note: tried to use across function for sd but got an error message i couldn't solve)

#function for SE 
SE <- function(x, type="normal") {
  if (type == "normal") {
    SE <- sd(x)/sqrt(length(x))
  }
  if (type == "poisson"){
    SE<- sqrt(mean(x)/length(x))
  }
  return(SE)
} #need to distinguish between distribution type in this function

s_se <- tibble(SE(s$height), SE(s$weight),SE(s$age), SE(s$zombies_killed, type = "poisson"), SE(s$years_of_education, type = "poisson"))#standard error for each variable 

#Confidence interval functions 
alpha = 0.05
CI.norm <- function(x){
  CI <- mean(x) + c(-1, 1) * qnorm(1-alpha/2) * SE(x) #se function - check there is one
  names(CI) <- c("Lower", "Upper")
  return(CI)
} #normal 

CI.sim <- function(x){
  sim <- NULL
  for (i in 1:1000){
    sim[i] <- mean(sample(x), length(x), replace = T)
  }
  CI <- quantile(sim, c(alpha/2, 1-alpha/2))
names(CI) <- c("Lower", "Upper")
return(CI)
} #simulation

#CI based on normal distribution
s_CI.norm <- rbind(CI.norm(s$height), CI.norm(s$weight), CI.norm(s$age), CI.norm(s$zombies_killed), CI.norm(s$years_of_education))

#CI based on simulation
s_CI.sim <- rbind(CI.sim(s$height), CI.sim(s$weight), CI.sim(s$age), CI.sim(s$zombies_killed), CI.sim(s$years_of_education))

single_samp_summary <- data.frame(Variable=c("Height", "Weight", "Age", "Kills", "Education"), MEAN = t(s_means), SD = t(s_sd), SE = t(s_se), CI.NORM = s_CI.norm, CI.SIMS = s_CI.sim)

```

---
Then draw another 99 random samples of 50 zombie apocalypse survivors out and calculate the mean for each of the these samples. Together with the first sample you drew out, you now have a set of 100 means for each variable (each based on 50 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of the sampling distribution for each variable?
---

```{r}
#99 random samples of 50 zombies 
set.seed(1)
k <- 99 # number of samples
n <- 50 # size of each sample
sample <- 1:k %>%
  map_dfr(
    ~ survivors %>% 
      slice_sample(n = n)%>%
      summarize(height_mean = mean(height), weight_mean =mean(weight), age_mean = mean(age), zombies_killed_mean = mean(zombies_killed), years_of_education_mean = mean(years_of_education))
  ) %>% mutate(n = 50)
sample

(samp_dist <- bind_rows(sample, s_means)) #combine single sample and sample of 99 into one df

samp_dist_mean <- samp_dist %>% summarise(height = mean(height_mean), weight = mean(weight_mean), age = mean(age_mean), kills = mean(zombies_killed_mean), edu = mean(years_of_education_mean)) #means for the sampling distribution 

#calculate mean of SE for variables from each sample
se_fun <- function(x){
  sd(x)/sqrt(100)
} #create a function for standard error
samp_dist_se <- samp_dist %>% summarise(height = se_fun(height_mean), weight = se_fun(weight_mean), age = se_fun(age_mean), kills = se_fun(zombies_killed_mean), edu = se_fun(years_of_education_mean)) #standard deviations (i.e. SE) for the sampling dist. 

comparison <- data.frame(Variables =c("Height", "Weight", "Age", "Kills", "Education"), "Mean_Samp_Dist" = t(samp_dist_mean), "SD_Samp_Dist" = t(samp_dist_se))
```

---
Finally, construct an 95% confidence interval for each mean directly from the sampling distribution of sample means using the central 95% that distribution (i.e., by setting the lower and upper CI bounds to 2.5% and 97.5% of the way through that distribution).
---

```{r}
#95% CI for each mean from the sampling dist. 
percent_ci <- 95
alpha <- 1 - percent_ci / 100
lower <- samp_dist_mean + qnorm(alpha / 2) * samp_dist_se
upper <- samp_dist_mean + qnorm(1 - alpha / 2) * samp_dist_se
ci <- c(lower, upper)

```
---
How do the standard deviations of the sampling distribution for each variable compare to the standard errors estimated from your first sample of size 50?
---
```{r}
compare <- data.frame(Variables=c("Height", "Weight", "Age", "Kills", "Education"), "Mean of Samp Dist" = t(samp_dist_mean), "SE from pop SD" = t(pop_sd/sqrt(n)), "Single Sample SE" = t(s_se), "Samp dist SE" = t(samp_dist_se))

#All are about the same. The SE for each variable is similar to the population estimate

```
---
What do sampling distributions for each variable mean look like? Are they normally distributed? 
---
```{r}

par(mfrow = c(1,2))
hist(sample$height_mean, main = "Height Means", xlab = "") #height
qqnorm(sample$height_mean, main = "")
qqline(sample$height_mean, col = "blue")
hist(sample$weight_mean, main = "Weight Means", xlab = "") #weight
qqnorm(sample$weight_mean, main = "")
qqline(sample$weight_mean, col = "blue")
hist(sample$age_mean, main = "Age Means", xlab = "") #age
qqnorm(sample$age_mean, main = "")
qqline(sample$age_mean, col = "blue")
hist(sample$zombies_killed_mean, main = "Zombie kills Means", xlab = "") #kills
qqnorm(sample$zombies_killed_mean, main = "")
qqline(sample$zombies_killed_mean, col = "blue")
hist(sample$years_of_education_mean, main = "Years of Education Means", xlab = "") #edu
qqnorm(sample$years_of_education_mean, main = "")
qqline(sample$years_of_education_mean, col = "blue")

```
---
What about for those variables that you concluded were not originally drawn from a normal distribution? 
All these distirbutions are normal, the two varibales that were previously poisson are now normal too. 
---
---
How do the two 95% CIs you estimated compare to one another (i.e., the CI based on one sample and its estimated standard deviation versus the CI based on simulation)?
---
```{r}
cbind(s_CI.norm, s_CI.sim) #based on normal distribution and based on simulation 

```
